### 2011-02-09 paper/experimentalSetup.tex
Fitness cleanup and rewording, shuffled some points between sections

\secref{platformDetails} discusses the hardware platform used for the
experiments presented in this paper, and \secref{fitnessEvaluation}
describes the method we used to measure a gait's fitness.

>>>>

An important aspect of this study was to evaluate gait generation and
learning algorithms in the noisy, unforgiving real world, instead of
using a simulator.  In this section we give details of the hardware
setup we used: \secref{platformDetails} discusses the robotic
platform, and \secref{fitnessEvaluation} describes the tracking system
we used to measure each gait's fitness.

### 2011-02-09 paper/fitnessEvaluation.tex
Fitness cleanup and rewording, shuffled some points between sections

As mentioned earlier, the metric for evaluating gaits was their
average speed.  In reality this was measured as the distance the robot
moved over a 12 second run, measured by the Wii remote tracking system
(\secref{platformDetails}).  The Euclidean distance between the
starting position and the ending position, in pixels, was reported
directly as the fitness (save for a tweak for HyperNEAT; see later in
this section).

Since only a single point on the robot --- the IR LED --- was measured
for the purposes of computing fitness, it was important that the
position of the IR LED accurately reflect the position of the robot as
a whole.  To enforce this constraint, the robot was always measured
while in the \emph{ready} position (see \figref{robot_close.jpg}).  If
this was not done, it would have resulted in giving extra fitness to,
for example, gaits that ended with the robot leaning into the
direction of travel.  While this might mean that a part of the robot
has traveled farther, this extra distance would not generalize to a
longer run.  Thus, we chose not to reward this behavior.

Wishing to measure start and end position in the same pose,

In order to ensure we could
start and end each run in the same position,

>>>>

\acmFigg{wiiMote_crop.jpg}{robot_led_crop.jpg}{1}{A Nintendo Wii
  remote provided the location of the robot by tracking the infrared
  LED mounted on the robot's antenna.  The LED position was measured
  in pixels and transmitted from the Wii remote to the robot via
  bluetooth.}

To track the position of the robot and thus determine gait fitness, we
mounted a Nintendo Wii remote on the ceiling and an infrared LED on
top of the robot (\figref{wiiMote_crop.jpg}).  The Wii remote contains
an IR camera that can track and report the position of up to four IR
sources in its image frame.  The resolution of the camera was 1024 by
768 pixels with view angles of about 40$^\circ$ by 30$^\circ$, which
produced a resolution of 1.7mm per pixel when mounted at a height of
2.63m.  At this height, the viewable window on the floor was
approximately 175 x 120 cm.

A separate tracking server, written in Python, ran on the robot and
interfaced with the Wii remote via bluetooth using the CWiid
library\cite{cwiid}.  Our fitness testing code communicated with this
server via socket connection and requested position updates at the
beginning and end of each run.

As mentioned earlier, the metric for evaluating gaits was their
average speed.  In reality this was measured as the distance the robot
moved over a 12 second run.  The Euclidean distance between the
starting position and the ending position, in pixels, was reported
directly as the fitness (save for a tweak for HyperNEAT; see below).

Since only a single point on the robot --- the IR LED --- was measured
for the purposes of computing fitness, it was important that the
position of the IR LED accurately reflect the position of the robot as
a whole.  To enforce this constraint, the robot was always measured
while in the \emph{ready} position (the position shown in
\figref{robot_close.jpg}).  This was done to prevent assigning extra
fitness to, for example, gaits that ended with the robot leaning into
the direction of travel.  While this would mean that a part of the
robot has traveled farther, this extra distance would not generalize
to a longer run.  Thus, we chose not to reward this behavior.

In order to measure start and end position in the same pose,

ensure we could
start and end each run in the same position,

### 2011-02-09 paper/platformDetails.tex
Fitness cleanup and rewording, shuffled some points between sections

The quadruped robot used in this study was assembled from parts
printed on the Objet Connex 500 3-D Printing System. It weights 1.88
kg with the on-board computer and measures approximately 38
centimeters from leg to opposite leg in the crouch position depicted
in \figref{robot_close.jpg}. The robot is actuated by 9 AX-12+
Dynamixel servos: one inner joint and one outer joint servo in each of
the four legs, and one servo at the center ``hip'' joint.  This final
unique servo allows the two halves of the robot to rotate with respect
to each other. \figref{topdown.png} shows the positions and numerical
designations of all nine servoes.

All of the computation for gait learning, fitness evaluation, and robot control
was performed on the compact on-board CompuLab Fit-PC2, running Ubuntu
Linux.  All gait generation, learning, and fitness evaluation code,
except HyperNEAT, is in Python and is available on our website
\cite{quadraWeb}.  HyperNEAT is written in C++.  We controlled the servos with the \code{pydynamixel} library \cite{pydynamixel}.  The robot
connects to a wireless network on boot, which enabled us to control it via SSH.

%\acmFigg{wiiMote.jpg}{robot_led.jpg}{1}{A Nintendo Wii remote tracks the
%  location of the robot, providing feedback about distance traveled,
%  in pixels, through an infrared LED mounted on top of the robot.}

\acmFigg{wiiMote_crop.jpg}{robot_led_crop.jpg}{1}{A Nintendo Wii remote tracks the
  location of the robot, providing feedback about distance traveled,
  in pixels, through an infrared LED mounted on top of the robot.}

To track the position of the robot and thus determine gait fitness, we
mounted a Nintendo Wii remote on the ceiling and an infrared LED on
top of the robot (\figref{wiiMote_crop.jpg}).  The Wii remote contains an
IR camera that can track and report the position of any IR sources in
its image frame.  The resolution of the camera was 1024 by 768 pixels,
which produced a resolution of 1.7mm per pixel when mounted at a
height of \edit{???}m.

A separate tracking server, written in Python, ran on the robot and
interfaced with the Wii remote via bluetooth using the CWiid
library\cite{cwiid}.  Our fitness testing code connected to this
server via socket and requested position updates at the beginning and
end of each run (see \secref{fitnessEvaluation}).

%to interface with the remote via bluetooth. A client then connects via
%a socket to the tracking server and requests position updates
%periodically. If the robot walks beyond the viewable range of the Wii
%remote, a prompt informed the user. The only human intervention
%required during the experiment was to move the robot back inside the
%viewable area and resume the run, and to handle any mechnanical
%failures that arose. This did not interrupt the learning process or
%result in the loss of data.

To run a gait on the robot, a \emph{gait
  function} is provided, any Python function that accepts a single input --- time
starting at 0 --- and outputs a list of nine commanded positions, one
for each servo.  To safeguard against limb collision with the robot
body, the robot control code cropped the commands to a safe range.
This range was [150, 770] for the inner leg servos, [30, 680] for the
outer leg servos, and [392, 623] for the center hip servo.

>>>>

The quadruped robot used in this study was assembled from parts
printed on the Objet Connex 500 3-D Printing System. It weights 1.88
kg with the on-board computer and measures approximately 38
centimeters from leg to opposite leg in the crouch position depicted
in \figref{robot_close.jpg}. The robot is actuated by 9 AX-12+
Dynamixel servos: one inner joint and one outer joint servo in each of
the four legs, and one servo at the center ``hip'' joint.  This final
unique servo allows the two halves of the robot to rotate with respect
to each other. \figref{topdown.png} shows the positions and numerical
designations of all nine servos.  Each servo could be commanded to a
position in the range [0,~1023], corresponding to a physical range
[-120$^{\circ}$,~+120$^{\circ}$].

All of the computation for gait learning, fitness evaluation, and
robot control was performed on the compact on-board CompuLab Fit-PC2,
running Ubuntu Linux.  All gait generation, learning, and fitness
evaluation code, except HyperNEAT, was written in Python and is
available on our website \cite{quadratotWeb}.  HyperNEAT is written in
C++.  We controlled the servos with the \code{pydynamixel} library
\cite{pydynamixel}.  The robot connected to a wireless network on
boot, which enabled us to control it via SSH.

To run a gait on the robot, one needed only to provide a \emph{gait
  function} --- any Python function that accepts a single input, time
starting at 0, and outputs a list of nine commanded positions, one for
each servo.  To safeguard against limb collision with the robot body,
the robot control code cropped the commands to a safe range.  This
range was [150,~770] for the inner leg servos, [30,~680] for the outer
leg servos, and [392,~623] for the center hip servo.

### 2011-02-09 paper/problemDefinition.tex
Fitness cleanup and rewording, shuffled some points between sections

\section{Problem definition}
% Precisely define the problem you are addressing (i.e. formally specify
% the inputs and outputs). Elaborate on why this is an interesting and
% important problem.
%
%\edit{write this.  Quadratot report section below}
%
%We are testing several different learning methods to design a
%parametrized gait for a quadruped robot from the Cornell Computational
%Synthesis Lab.
%
%The output each of the learning algorithms is a function of time,
%$g(t)$, that outputs a vector of commanded motor positions.  This
%function is generated using a parametrized motion model, described in
%\secref{implement}.
%
%The robot executes these commands and measures its change in location
%using the tracking system described in \secref{implement}.  The input
%to the learning algorithms is this measured displacement, which the
%algorithms attempt to maximize. This displacement is measured for each
%gait over a constant length run, usually 12 seconds.
%
%A comparison and evaluation of the many different methods available
%for optimizing the gait of legged robots will be useful for future
%work on this challenging multidimensional control problem.

@@

The ultimate goal was to design gaits that were as fast as possible,
so to optimize for speed, our performance metric was displacement over
a period of motion lasting approximately 10 seconds.  For details of
exactly how this displacement was measured, see
\secref{fitnessEvaluation}.

>>>>

\section{Problem definition}
\seclabel{problemDefinition}

% Precisely define the problem you are addressing (i.e. formally specify
% the inputs and outputs). Elaborate on why this is an interesting and
% important problem.
%
%\edit{write this.  Quadratot report section below}
%
%We are testing several different learning methods to design a
%parametrized gait for a quadruped robot from the Cornell Computational
%Synthesis Lab.
%
%The output each of the learning algorithms is a function of time,
%$g(t)$, that outputs a vector of commanded motor positions.  This
%function is generated using a parametrized motion model, described in
%\secref{implement}.
%
%The robot executes these commands and measures its change in location
%using the tracking system described in \secref{implement}.  The input
%to the learning algorithms is this measured displacement, which the
%algorithms attempt to maximize. This displacement is measured for each
%gait over a constant length run, usually 12 seconds.
%
%A comparison and evaluation of the many different methods available
%for optimizing the gait of legged robots will be useful for future
%work on this challenging multidimensional control problem.

@@

The ultimate goal was to design gaits that were as fast as possible,
so to optimize for speed, our performance metric was displacement over
a period of motion lasting approximately 12 seconds.  For details of
exactly how this displacement was measured, see
\secref{fitnessEvaluation}.

### 2011-02-09 paper/results.tex
Added plots
Fitness cleanup and rewording, shuffled some points between sections

@@

>>>>

\edit{maybe useful vvvvv}

To be as fair as possible when obtaining the results presented in
\secref{results}, we picked three random points --- $\vec{\theta}_A$,
$\vec{\theta}_B$, and $\vec{\theta}_C$ --- and started all seven
methods at these points...

Each of the parameter ??? methods was run on 3 different initial
parameter vectors, in order to fairly compare the algorithms. We allowed 
each run for the parameter ??? methods to
continue until the results plateaued (no improvement for one third of
the policies seen so far). Three runs of HyperNEAT were completed,
each with a different initial seed and run for 20 generations.

\edit{$\wedge\wedge\wedge\wedge\wedge\wedge\wedge\wedge$}

@@

\acmFig{neat_110115_211410_00000_002_filt}{1}{Caption here...???}
\acmFig{neat_110115_211410_00000_002_filt_zoom}{1}{Caption here...???}
